Presolving for\\ Mixed-Integer Semidefinite Optimization

### 2020-12-10 tex/MISDP-Presolving/MISDPpresolving.tex
fixed description of SOC formulation as SDP
add twominorprodconss and twominorsocconss to presolving document

The following inequalities are known from the literature, see
Mars~\cite{Mar13} and  Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
\eqref{MISDP}:
\begin{itemize}
\item Any feasible $y \in \R^m$ satisfies $A(y) \succeq 0$, which implies
  that the diagonal entries of $A(y)$ are nonnegative, i.e., for all
  $i \in [n]$:
  \[
    \sum_{k = 1}^m (A^k)_{ii}\, y_k - (A^0)_{ii} \geq 0.
  \]
\item If for some $i$, $j \in [n]$, $A^0_{ij} \neq 0$, and $A^0_{ii} = 0$,
  $A^k_{ij} = 0$ for all $k \in [m]$, $A^k_{ii} = 0$ for all continuous
  variables and $\ell_i \geq 0$ for all integer variables,
  the following inequality is valid:
  \[
    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq 1.
  \]
  Note that the conditions imply that $A^0$ is not psd.
\item For all $X \succeq 0$, we have
  \[
    X_{ii} + X_{jj} - 2\, X_{ij} \geq 0,\qquad
    X_{ii} + X_{jj} + 2\, X_{ij} \geq 0.
  \]
  This follows by using a 2 by 2 minor and multiplying from left
  and right by the all-ones vector and $[1,-1]$, respectively. It also
  follows from~\eqref{eq:TwoByTwoMinorInequality}.
\item The previous inequalities translated to the matrix pencil are:
  \begin{align*}
    & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
    \quad\Leftrightarrow\quad&
    \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} - 2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
  \end{align*}
\end{itemize}

@@

\begin{remark}
  Note that in the pure form, the propagated inequalities from
  Lemma~\ref{lem:Propagation} dominate the inequalities given in
  Section~\ref{sec:ImpliedLinEqs}:
  \begin{align*}
    X_{st} \leq \sqrt{U_{ss} \cdot U_{tt}} \leq \frac{1}{2}\big( U_{ss} + U_{tt}\big),
  \end{align*}
  because of the inequality between the arithmetic and geometric mean. Does
  this also hold in the general case of a (dual) LMI?
\end{remark}

@@

Indeed, SOCP is special case of SDP, because:
\[
  (x,t) \in L_{n+1} \quad\Leftrightarrow\quad
  \tilde{X} \define
  \begin{pmatrix}
    t & x\T\\
    x & I_n
  \end{pmatrix}
  \succeq 0.
\]
To see this, we need the following result adapted
from~\cite[Section~A.5.5]{BoyV09}. Let
\[
  D = \begin{pmatrix}
    A & B\T\\
    B & C
  \end{pmatrix}
\]
with a positive definite $C \succ 0$. Then $D \succeq 0$ if and only if
$S = A - B\T C^{-1} B \succeq 0$ and $A \succeq 0$\footnote{Note that the
  condition $A \succeq 0$ is missing in~\cite{BoyV09}.}. For $D = \tilde{X}$, we
get: $\tilde{X} \succeq 0$ if and only if $t \geq 0$, $t - x\T x \geq
0$. The latter is equivalent to $(x,t) \in L_{n+1}$.

>>>>

The following inequalities are known from the literature, see
Mars~\cite{Mar13} and  Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
\eqref{MISDP}:
\begin{itemize}
\item Any feasible $y \in \R^m$ satisfies $A(y) \succeq 0$, which implies
  that the diagonal entries of $A(y)$ are nonnegative, i.e., for all
  $i \in [n]$:
  \begin{align}\label{eq:diaggezero}
    \sum_{k = 1}^m (A^k)_{ii}\, y_k - (A^0)_{ii} \geq 0.
  \end{align}
\item If for some $i$, $j \in [n]$, $A^0_{ij} \neq 0$, and $A^0_{ii} = 0$,
  $A^k_{ij} = 0$ for all $k \in [m]$, $A^k_{ii} = 0$ for all continuous
  variables and $\ell_i \geq 0$ for all integer variables,
  the following inequality is valid:
  \begin{align}\label{eq:diagzeroimpl}
    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq1.
  \end{align}
  Note that the conditions imply that $A^0$ is not psd.
\item For all $X \succeq 0$, we have
  \begin{align}\label{eq:2minorlinP}
    X_{ii} + X_{jj} - 2\, X_{ij} \geq 0,\qquad
    X_{ii} + X_{jj} + 2\, X_{ij} \geq 0.
  \end{align}
  This follows by using a 2 by 2 minor and multiplying from left
  and right by the all-ones vector and $[1,-1]$, respectively. It also
  follows from~\eqref{eq:TwoByTwoMinorInequality}.
\item The previous inequalities translated to the matrix pencil are:
  \begin{equation}\label{eq:2minorlinD}
    \begin{aligned}
      & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
      \quad\Leftrightarrow\quad& \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} -
      2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
    \end{aligned}
  \end{equation}
\item If $A^k_{ii} = A^k_{jj} = 0$ for all~$k \in [m]$
  and~$A^0_{ii}A^0_{jj} \geq 0$ for some $i \neq j \in [n]$, then the
  following inequality is valid:
  \begin{align}\label{eq:2minorprodD}
    \sum_{k=1}^m A^k_{ij} y_k \geq A^0 - \sqrt{A^0_{ii}A^0_{jj}}.
  \end{align}
  Furhtermore, if exactly one~$A^k_{ij} \neq 0$, then this yields upper or
  lower bounds for the corresponding variable~$y_k$, depending on the sign
  of~$A^k_{ij}$.
\item For a positive semidefinite matrix $X$ the following quadratic
  inequality holds by using the 2 by 2 minor:
  $X_{st}^2 \leq X_{ss} X_{tt}$. This constraint is SOC representable:
  \begin{align*}
    \left\|\binom{2\,X_{st}}{X_{ss} - X_{tt}}\right\| &\leq X_{ss} +
                                                        X_{tt}. \\
    \quad\Leftrightarrow\quad 4\, X_{st}^2 + (X_{ss}^2 - 2\,X_{ss}\, X_{tt}
    + X_{tt}^2) & \leq X_{ss}^2 + 2\, X_{ss}\,X_{tt} + X_{tt}^2.
  \end{align*}
  Translated to the matrix pencil notation the cut looks as follows:
  \begin{equation}\label{eq:2minorSOC}
    \begin{aligned}
      &\left\|\binom{2\, \sum_{i=1}^m (A_i)_{st}\, y_i - 2\, (A_0)_{st}}{ q
          \sum_{i=1}^m (A_i)_{ss}\, y_i - (A_0)_{ss} - \sum_{i=1}^m
          (A_i)_{tt}\, y_i + (A_0)_{tt}}\right\| \\ &\leq \sum_{i=1}^m
      (A_i)_{ss}\, y_i - (A_0)_{ss} + \sum_{i=1}^m (A_i)_{tt}\, y_i -
      (A_0)_{tt}.
    \end{aligned}
  \end{equation}
  These constraints can be added directly as SOC-constraints.
\end{itemize}

@@

\begin{remark}
  Note that in the pure form, the propagated inequalities from
  Lemma~\ref{lem:Propagation} dominate the inequalities
  in~\eqref{eq:2minorlinP}:
  \begin{align*}
    X_{st} \leq \sqrt{U_{ss} \cdot U_{tt}} \leq \frac{1}{2}\big( U_{ss} + U_{tt}\big),
  \end{align*}
  because of the inequality between the arithmetic and geometric mean. Does
  this also hold in the general case of a (dual) LMI?
\end{remark}

@@

Indeed, SOCP is special case of SDP, because:
\[
  (x,\sqrt{t}) \in L_{n+1} \quad\Leftrightarrow\quad
  \tilde{X} \define
  \begin{pmatrix}
    t & x\T\\
    x & I_n
  \end{pmatrix}
  \succeq 0.
\]
To see this, we need the following result adapted
from~\cite[Section~A.5.5]{BoyV09}. Let
\[
  D = \begin{pmatrix}
    A & B\T\\
    B & C
  \end{pmatrix}
\]
with a positive definite $C \succ 0$. Then $D \succeq 0$ if and only if
$S = A - B\T C^{-1} B \succeq 0$ and $A \succeq 0$\footnote{Note that the
  condition $A \succeq 0$ is missing in~\cite{BoyV09}.}. For $D = \tilde{X}$, we
get: $\tilde{X} \succeq 0$ if and only if $t \geq 0$, $t - x\T x \geq
0$. The latter is equivalent to $(x,\sqrt{t}) \in L_{n+1}$.
